---
title: "Table 1: sPDL1 medians"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clear the environment
rm(list = ls())
# Free up memory by forcing garbage collection
invisible(gc())  
# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9
##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

# Paths, Packages, Functions, Palettes

```{r Paths}

# Provide paths
data_dir <- "./Data"
results_dir <- "./Outputs/Manuscript"

```


```{r LoadPackages}
library(knitr) # report generation

##_ alternatively load tidyverse which bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
##_ use suppressPackageStartup() to hide package load messages
suppressPackageStartupMessages(library(tidyverse))  
```

\newpage
# Objective

Table of median sPDL1 values to refer to throughout paper.

## Data Sources

```{r input, include= FALSE}

#DFCI results reformatted and clinical annotation by Petra RM
#this file does not have BOR
#tidydata_file <- paste(data_dir,"DFCI_sPDL1_TidyData2.txt", sep = "/")


#read annotated dataset from data_dir
dataset_rds <- paste(data_dir,
				   "sPDL1_CM9_CM38_Annotated_v2.Rds", sep = "/")

data <- readRDS(dataset_rds)

# all NE in CM38 Part 1 can be used as PD
data$BOR[which(data$BOR == "NE" & data$Dataset == "CM38")] <- "PD"
data$BOR3[which(data$BOR3 == "NE" & data$Dataset == "CM38")] <- "PD"
data$OR_group[which(is.na(data$OR_group) & data$Dataset == "CM38")] <- "NonResponder"
data$OR_num[which(is.na(data$OR_num) & data$Dataset == "CM38")] <- "0"


```


Data plus harmonized annotation for CM9 and CM38 (and ultimately its codebook), was read from github:

+ *`r tidydata_file`*
+ *`r dataset_rds`*

All NE in CM38 Part 1 can be used as PD. This was amended.

\newpage
# Results

### CM9 Medians

Final table summarizes the median sPDL1 values for whole cohort, for 2L, and for each BOR3 group at each timepoint. 

ULOQ samples were removed before reporting medians for the change over time.

```{r cm9_tidydata}

data_9 <- filter(data,
               Dataset == "CM9")

#make a table of one row per spdl1 observation with annotation

tidydata_9 <- data_9%>%
	gather(key =  "Day",
		   value = "sPDL1_pgml",
		  PDL1_Av_Day1,
		  PDL1_Av_Day29,
		  PDL1_Av_Day63)%>%
	select(USUBJID,
		   Dataset,
		   PriorTherapyGroup,
		   BOR3,
		   OR_group,
		   TPScat,
		   Day,
		   sPDL1_pgml,
		   ULOQ_FLAG_Day1,
		   ULOQ_FLAG_Day29,
		   ULOQ_FLAG_Day63)


#Make Day into an ordered factor
tidydata_9$Day <- factor(tidydata_9$Day, 
							levels = c("PDL1_Av_Day1",
									   "PDL1_Av_Day29",
									   "PDL1_Av_Day63"))


#Make PriorTherapyGroup into an ordered factor
tidydata_9$PriorTherapyGroup <- factor(tidydata_9$PriorTherapyGroup, 
							levels = c("Naive",
									   "Pretreated"))

```

I derived a table of one row per sPDL1 observation with CM9 annotation. 

+ to make a summary table


```{r cm9_day_by_BOR3}

# calculate day medians in CM9 spdl1 dataset

day <- tidydata_9 %>% 
	group_by(Dataset,Day) %>% 
  summarise(Day1_N = n(),
  		  Median = round(median(sPDL1_pgml, na.rm = TRUE), 0))


day$Day <- gsub(pattern = "PDL1_Av_",
			   x = day$Day,  
			   replacement = "")
	
day <-spread(day, 
				Day,
				Median)

day <- rename(Cohort = Dataset, day)

# calculate day medians in 2L CM9 spdl1 dataset

day_2L <- tidydata_9 %>%
  filter(PriorTherapyGroup == "Pretreated")%>%
	group_by(PriorTherapyGroup,Day) %>% 
  summarise(Day1_N = n(),
  		  Median = round(median(sPDL1_pgml, na.rm = TRUE), 0))


day_2L$Day <- gsub(pattern = "PDL1_Av_",
			   x = day_2L$Day,  
			   replacement = "")
	
day_2L <-spread(day_2L, 
				Day,
				Median)

day_2L <- rename(Cohort = PriorTherapyGroup, day_2L)

# calculate day/BOR medians in CM9 spdl1 dataset

daybor <- tidydata_9 %>% 
	filter(BOR3 != "NE")%>%
	group_by(BOR3,Day) %>% 
  summarise(Day1_N = n(),
  		  Median = round(median(sPDL1_pgml, na.rm = TRUE), 0))


daybor$Day <- gsub(pattern = "PDL1_Av_",
			   x = daybor$Day,  
			   replacement = "")
	
daybor <-spread(daybor, 
				Day,
				Median)

daybor <- rename(Cohort = BOR3, daybor)

#Put the 3 summaries together
day_medians <- rbind(day,
                     day_2L,
                     daybor) 

#annotate the trial on the left
day_medians$Trial <- "CM-009"
day_medians <- day_medians%>%
  select(Trial, everything())

#harmonize the day63 column
day_medians <- rename(Day43or63 = Day63, day_medians)
```


```{r cm9_change_by_BOR3}

diff_medians <- NULL

#calculate diff medians for whole cohort, without ULOQ data

day29diff_all <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day29 == "0")%>%
	group_by(Dataset) %>% 
  summarise(Day29_Pairs = n(),
             Day1toDay29Change = round(median(PDL1_DIFF_Day29, na.rm = TRUE), 0))

day29diff_all <- rename(Cohort = Dataset, day29diff_all)

day63diff_all <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day63 == "0")%>%
	group_by(Dataset) %>% 
  summarise(Day63_Pairs = n(),
             Day1toDay63Change = round(median(PDL1_DIFF_Day63, na.rm = TRUE), 0))

day63diff_all <- rename(Cohort = Dataset, day63diff_all)

diff_all <- day29diff_all%>%
  left_join(day63diff_all)

#calculate diff medians for 2L cohort, without ULOQ data

day29diff_2L <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day29 == "0",
		   PriorTherapyGroup == "Pretreated")%>%
	group_by(PriorTherapyGroup) %>% 
  summarise(Day29_Pairs = n(),
             Day1toDay29Change = round(median(PDL1_DIFF_Day29, na.rm = TRUE), 0))

day29diff_2L <- rename(Cohort = PriorTherapyGroup, day29diff_2L)

day63diff_2L <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day63 == "0",
		   PriorTherapyGroup == "Pretreated")%>%
	group_by(PriorTherapyGroup) %>% 
  summarise(Day63_Pairs = n(),
             Day1toDay63Change = round(median(PDL1_DIFF_Day63, na.rm = TRUE), 0))

day63diff_2L <- rename(Cohort = PriorTherapyGroup, day63diff_2L)

diff_2L <- day29diff_2L%>%
  left_join(day63diff_2L)


#calculate diff medians for bor, without ULOQ data

day29diff_bor <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day29 == "0"
		   &BOR3 != "NE")%>%
	group_by(BOR3) %>% 
  summarise(Day29_Pairs = n(),
             Day1toDay29Change = round(median(PDL1_DIFF_Day29, na.rm = TRUE), 0))

day29diff_bor <- rename(Cohort = BOR3, day29diff_bor)

day63diff_bor <- data_9 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day63 == "0"
		   &BOR3 != "NE")%>%
	group_by(BOR3) %>% 
  summarise(Day63_Pairs = n(),
             Day1toDay63Change = round(median(PDL1_DIFF_Day63, na.rm = TRUE), 0))

day63diff_bor <- rename(Cohort = BOR3, day63diff_bor)

diff_bor <- day29diff_bor%>%
  left_join(day63diff_bor)


#Put the 3 summaries together
diff_medians <- rbind(diff_all,
                     diff_2L,
                     diff_bor) 

#harmonize the day63 columns
diff_medians <- rename(`Day 43or63 Pairs` = Day63_Pairs, diff_medians)
diff_medians <- rename(`Day1 to Day43or63 Change` = Day1toDay63Change, diff_medians)


```

```{r cm9_table}

#Unite the tables

medians_cm9 <- left_join(day_medians,
					diff_medians,
					by = "Cohort")

#print the table

kable(medians_cm9, digits = 1,
      caption = "CM9: Median values for sPDL1, pg/ml")
```


### CM38 Medians

Final table summarizes the median sPDL1 values for whole cohort,and for each BOR3 group at each timepoint. 

ULOQ samples were removed before reporting medians for the change over time.

```{r cm38_tidydata}

data_38 <- filter(data,
               Dataset == "CM38")

#make a table of one row per spdl1 observation with annotation

tidydata_38 <- data_38%>%
	gather(key =  "Day",
		   value = "sPDL1_pgml",
		  PDL1_Av_Day1,
		  PDL1_Av_Day29,
		  PDL1_Av_Day43)%>%
	select(USUBJID,
		   Dataset,
		   PriorTherapyGroup,
		   BOR3,
		   OR_group,
		   TPScat,
		   Day,
		   sPDL1_pgml,
		   ULOQ_FLAG_Day1,
		   ULOQ_FLAG_Day29,
		   ULOQ_FLAG_Day43)


#Make Day into an ordered factor
tidydata_38$Day <- factor(tidydata_38$Day, 
							levels = c("PDL1_Av_Day1",
									   "PDL1_Av_Day29",
									   "PDL1_Av_Day43"))


```

I derived a table of one row per sPDL1 observation with CM9 annotation. 

+ to make a summary table


```{r cm38_day_by_BOR3}

# calculate day medians in CM38 spdl1 dataset

day <- tidydata_38 %>% 
	group_by(Dataset,Day) %>% 
  summarise(Day1_N = n(),
  		  Median = round(median(sPDL1_pgml, na.rm = TRUE), 0))


day$Day <- gsub(pattern = "PDL1_Av_",
			   x = day$Day,  
			   replacement = "")
	
day <-spread(day, 
				Day,
				Median)

day <- rename(Cohort = Dataset, day)

# calculate day/BOR medians in CM38 spdl1 dataset

daybor <- tidydata_38 %>% 
	filter(BOR3 != "NE")%>%
	group_by(BOR3,Day) %>% 
  summarise(Day1_N = n(),
  		  Median = round(median(sPDL1_pgml, na.rm = TRUE), 0))


daybor$Day <- gsub(pattern = "PDL1_Av_",
			   x = daybor$Day,  
			   replacement = "")
	
daybor <-spread(daybor, 
				Day,
				Median)

daybor <- rename(Cohort = BOR3, daybor)

#Put the 3 summaries together
day_medians <- rbind(day,
                     daybor) 

#annotate the trial on the left
day_medians$Trial <- "CM-038"
day_medians <- day_medians%>%
  select(Trial, everything())

#harmonize the day63 column
day_medians <- rename(Day43or63 = Day43, day_medians)
```


```{r cm38_change_by_BOR3}

diff_medians <- NULL

#calculate diff medians for whole cohort, without ULOQ data

day29diff_all <- data_38 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day29 == "0")%>%
	group_by(Dataset) %>% 
  summarise(Day29_Pairs = n(),
             Day1toDay29Change = round(median(PDL1_DIFF_Day29, na.rm = TRUE), 0))

day29diff_all <- rename(Cohort = Dataset, day29diff_all)

day43diff_all <- data_38 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day43 == "0")%>%
	group_by(Dataset) %>% 
  summarise(Day43_Pairs = n(),
             Day1toDay43Change = round(median(PDL1_DIFF_Day43, na.rm = TRUE), 0))

day43diff_all <- rename(Cohort = Dataset, day43diff_all)

diff_all <- day29diff_all%>%
  left_join(day43diff_all)

#calculate diff medians for bor, without ULOQ data

day29diff_bor <- data_38 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day29 == "0"
		   &BOR3 != "NE")%>%
	group_by(BOR3) %>% 
  summarise(Day29_Pairs = n(),
             Day1toDay29Change = round(median(PDL1_DIFF_Day29, na.rm = TRUE), 0))

day29diff_bor <- rename(Cohort = BOR3, day29diff_bor)

day43diff_bor <- data_38 %>% 
	filter(ULOQ_FLAG_Day1 == "0"
		   & ULOQ_FLAG_Day43 == "0"
		   &BOR3 != "NE")%>%
	group_by(BOR3) %>% 
  summarise(Day43_Pairs = n(),
             Day1toDay43Change = round(median(PDL1_DIFF_Day43, na.rm = TRUE), 0))

day43diff_bor <- rename(Cohort = BOR3, day43diff_bor)

diff_bor <- day29diff_bor%>%
  left_join(day43diff_bor)


#Put the 3 summaries together
diff_medians <- rbind(diff_all,
                     diff_bor) 

#harmonize the day43 columns
diff_medians <- rename(`Day 43or63 Pairs` = Day43_Pairs, diff_medians)
diff_medians <- rename(`Day1 to Day43or63 Change` = Day1toDay43Change, diff_medians)


```

```{r cm38_table}

#Unite the tables

medians_cm38 <- left_join(day_medians,
					diff_medians,
					by = "Cohort")

#print the table

kable(medians_cm38, digits = 1,
      caption = "CM38: Median values for sPDL1, pg/ml")
```

```{r unite_summary}

medians_cm9_cm38 <- rbind(medians_cm9,
                          medians_cm38)

#better colnames
medians_cm9_cm38 <-rename(`Day1 to Day29 Change` = Day1toDay29Change, medians_cm9_cm38)
medians_cm9_cm38 <-rename(`Day1 (n=)` = Day1_N, medians_cm9_cm38)
medians_cm9_cm38 <-rename(`Day29 (Pairs, n=)` = Day29_Pairs, medians_cm9_cm38)
medians_cm9_cm38 <-rename(`Day 43 or 63 (Pairs, n=)` = `Day 43or63 Pairs`, medians_cm9_cm38)

#Better variables
medians_cm9_cm38$Cohort <- gsub("CM9","Trial Cohort",medians_cm9_cm38$Cohort)
medians_cm9_cm38$Cohort <- gsub("CM38","Trial Cohort",medians_cm9_cm38$Cohort)
medians_cm9_cm38$Cohort <- gsub("Pretreated","2L Cohort",medians_cm9_cm38$Cohort)

```


\newpage
# Outputs

```{r output}

medians_file <- paste(results_dir,"TABLE_T1_Cohort_Medians_sPDL1.txt",sep = "/" )

write_tsv(medians_cm9_cm38,medians_file)

```


Assembled table was written to github:

+ *`r medians_file`*