---
title: "Figure 1: sPDL1 versus Clinical Characteristics v2"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clear the environment
rm(list = ls())
# Free up memory by forcing garbage collection
invisible(gc())  
# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9
##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

# Paths, Packages, Functions, Palettes

```{r Paths}

# Provide paths
data_dir <- "./Data"
results_dir <- "./Outputs/Manuscript"

```


```{r LoadPackages}
library(knitr) # report generation
library(ggpubr)

#library(reshape2)#melt

library(RColorBrewer)
library(cowplot)

##_ alternatively load tidyverse which bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
##_ use suppressPackageStartup() to hide package load messages
suppressPackageStartupMessages(library(tidyverse))  
```

```{r functions}

source("./Code/Code_PostBMS/theme_dj.R")

```

```{r Palettes}

# color map for response
flagColors <- c("1" = "black", "0" = "grey")

# shape map for flag, use stroke = 2 to thicken line
flagshape_empty <- c("1" = 0, 
               "0" = 1)

```

\newpage
# Objective

Figure 1: My interpretation

1. 1a: CM9 spdl1 by timepoint
1. 1b: CM9 spdl1 by prior
1. 1c: CM9 spdl1 by TPS
1. 1d: CM38 spdl1 by timepoint
1. 1e: CM38 spdl1 by prior
1. 1f: CM38 spdl1 by TPS

Supplemental about clinical
1. CM9 spdl1 vs PBRM1, VHL1, age,sex
1. CM38 spdl1 vs subtype,age,sex

Supplemental about TPS?
1. Supplemental . Melanoma TPS not significantly associated with responders, but TPS significantly higher in CR/PR than PD. 
1. Supplemental . No signif difference in RCC TPS for subjects with PD, SD, CR/CP 




## Data Sources


```{r input, include= FALSE}

#DFCI results, formatted and annotated by Petra R-M

#read annotated dataset from data_dir

dataset_rds <- paste(data_dir,
				   "sPDL1_CM9_CM38_Annotated_v2.Rds", sep = "/")

data <- readRDS(dataset_rds)

# all NE in CM38 Part 1 can be used as PD
data$BOR[which(data$BOR == "NE" & data$Dataset == "CM38")] <- "PD"
data$BOR3[which(data$BOR3 == "NE" & data$Dataset == "CM38")] <- "PD"
data$OR_group[which(is.na(data$OR_group) & data$Dataset == "CM38")] <- "NonResponder"
data$OR_num[which(is.na(data$OR_num) & data$Dataset == "CM38")] <- "0"

#Make ULOQ flag a factor for coloring 
data$ULOQ_FLAG_Day1 <- factor(data$ULOQ_FLAG_Day1 ,
                              levels = c("0",
                                         "1"))

#read annotated tidy dataset from data_dir
#created in CM009_CM038_Make_tidydata.Rmd
#this has all NE in CM38 Part 1 as PD
tidydataset_rds <- paste(data_dir,
				   "sPDL1_CM9_CM38_tidydata_clin.Rds", sep = "/")

tidydata <- readRDS(tidydataset_rds)

## CM9  clin annotation
cm9_txt <- paste(data_dir,"CA209-009_Subject_Sample Annotation.txt", sep = "/")

cm9 <- read_tsv(cm9_txt)

#CM38  clin annotation
cm38_rds <- paste(data_dir,"CA209038_FA01_PetraRM_v2.Rds", sep = "/")

cm38 <- readRDS(cm38_rds)


```


Data plus harmonized annotation for CM9 and CM38 (and ultimately its codebook), was read from github:

+ *`r dataset_rds`*
+ *`r tidydataset_rds`*

Further Clinical annotation was read from:

+ *`r cm9_txt`*
+ *`r cm38_rds`*

```{r cm9_clinical}

data9 <- data%>%
  filter(Dataset == "CM9")

#need to merge in PBRM1 status, sex, age

#CM9 add genomic annotation

cm9columns <- c("USUBJID",
                "SEX",
                "AGE",
                "PBRM1","VHL","SETD2","BAP1")

data9 <- data9 %>%
          left_join(select(cm9,
                           one_of(cm9columns)),
                           by = c("USUBJID" = "USUBJID"))

#CLASSIFY PBRM1
data9$PBRM1_class <- NA
data9$PBRM1_class <- "Mutant"
data9$PBRM1_class[which(data9$PBRM1 == "WT")] <- "WT"
data9$PBRM1_class[is.na(data9$PBRM1)] <- "ND"
data9$PBRM1_class <- factor(data9$PBRM1_class, levels = c("Mutant", "WT", "ND"))


#CLASSIFY VHL
data9$VHL_class <- NA
data9$VHL_class <- "Mutant"
data9$VHL_class[which(data9$VHL == "WT")] <- "WT"
data9$VHL_class[is.na(data9$VHL)] <- "ND"
data9$VHL_class <- factor(data9$VHL_class, levels = c("Mutant", "WT", "ND"))

#factorize sex
data9$SEX <- factor(data9$SEX, levels = c("1", "2"))

#Dichotomize Age
data9$age_class <- NA
data9$age_class <- "< 65"
data9$age_class[which(data9$AGE > 64)] <- ">= 65"
data9$age_class <- factor(data9$age_class, levels = c("< 65",  ">= 65"))

```

These annotations were added to the CM9 data:

+ *`r cm9columns`*

```{r cm38_clinical}

data38 <- data%>%
  filter(Dataset == "CM38")

#need to merge in subtype, sex, age

#CM38 add  annotation

cm38columns <- c("USUBJID",
                "SEX",
                "AGE",
                "AGEGR2",
                "MELSUBT",
                "CNSMET",
                "BRAFSTA",
                "LDHCAT1",
                "LDHCAT2",
                "SMKNG")

data38 <- data38 %>%
          left_join(select(cm38,
                           one_of(cm38columns)),
                           by = c("USUBJID" = "USUBJID"))


#factorize sex
data38$SEX <- factor(data38$SEX, levels = c("M", "F"))

#factorize Age
cm38$AGEGR2 <- factor(cm38$AGEGR2, levels = c("< 65",  ">= 65"))

#factorize cm38$MELSUBT
cm38$MELSUBT<- factor(cm38$MELSUBT,
                      levels = c("CUTANEOUS",
                                 "MUCOSAL",
                                 "OCULAR/UVEAL",
                                 "ACRAL",
                                 "OTHER"))
```

These annotations were added to the CM38 data:

+ *`r cm38columns`*


```{r tidydata9}

#make a table of one row per spdl1 observation with annotation

tidydata_9 <- tidydata%>%
  filter(Dataset == "CM9")


#Make Day neater
tidydata_9$Day <- gsub("Day","Day ", tidydata_9$Day)
tidydata_9$Day <- gsub("Day 1","Baseline", tidydata_9$Day)


#Make Day into an ordered factor
tidydata_9$Day <- factor(tidydata_9$Day, 
							levels = c("Baseline",
									   "Day 29",
									   "Day 63"))

```

I derived a table of one row per sPDL1 observation with CM9 annotation. 

+ to make a summary table
+ For plots of change by patient

```{r tidydata38}

#make a table of one row per spdl1 observation with annotation
tidydata_38 <- tidydata%>%
  filter(Dataset == "CM38")

#Make Day neater
tidydata_38$Day <- gsub("Day","Day ", tidydata_38$Day)
tidydata_38$Day <- gsub("Day 1","Baseline", tidydata_38$Day)


#Make Day into an ordered factor
tidydata_38$Day <- factor(tidydata_38$Day, 
							levels = c("Baseline",
									   "Day 29",
									   "Day 43"))

```

I filtered a table of one row per sPDL1 observation with CM38 annotation. 

+ to make a summary table
+ For plots of change by patient



\newpage
# Results

## Figure 1 a: sPDL1 versus Timepoint in CM9

```{r Time_paired_boxplot_cm9}

#this tidy data already has null observations removed
no_ne <- tidydata_9

# Visualize: Specify the comparisons I want
my_comparisons <- list( c("Baseline", "Day 29"),
                        c("Baseline", "Day 63"))

#count samples in data
plotcount <- nrow(no_ne)

#Count the samples
count_baseline <- sum(no_ne$Day == "Baseline")
count_day29 <- sum(no_ne$Day == "Day 29")
count_day63 <- sum(no_ne$Day == "Day 63" )

spdl1_byDay_9 <- no_ne %>%
  ggplot(aes(x = Day, y = sPDL1)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQflag, shape = ULOQflag),
            			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
  theme_dj(7) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: Time",
       y = "sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
      	scale_x_discrete(labels=c("Baseline" = paste0("Baseline\nN=",count_baseline), 
							  "Day 29" = paste0("Day 29\nN=",count_day29),
							  "Day 63" = paste0("Day 63\nN=",count_day63)))+
  	stat_compare_means(method="wilcox.test", size = 3,
					   aes(label = paste0("P = ", ..p.format..)),
					   comparisons = my_comparisons,
					   vjust = "inward")+
theme(legend.position = "none")

  
spdl1_byDay_9
```

## Figure 1 b: Baseline sPDL1 versus Prior Therapy CM9

```{r wilcoxon_day1_PriorTherapyGroup_CM9}

# Get CM9 baseline data

print(wilcox.test(PDL1_Av_Day1 ~ PriorTherapyGroup, data = data9))
```

CM9: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the two PriorTherapyGroup categories are/are NOT identical populations.

```{r boxplot_baseline_prior_9}

# Specify desired comparison
my_comparisons <- list( c("Pretreated", "Naive"))

# Count number of datapoints plotted
plotcount <- nrow(data9)

#Count the samples
count_naive <- sum(data9$PriorTherapyGroup == "Naive")
count_pretreated <- sum(data9$PriorTherapyGroup == "Pretreated")

spdl1_vs_prior_plot_9 <- ggplot(data9,
                          aes(y = PDL1_Av_Day1, x = PriorTherapyGroup)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
            			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: Prior VEGFi status",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("Naive" =  paste0("Naive\nN=",count_naive), 
							  "Pretreated" =  paste0("Experienced\nN=",count_pretreated)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_prior_plot_9

```

## Figure 1 c: Baseline sPDL1 versus Tumor PDL1 Score in CM9

```{r spdl1_vs_TPScat_plot_9}

# Specify desired comparison
my_comparisons <- list( c("0%", "1to5%"), 
#                        c("1to5%", "5+%"), 
                        c("0%", "5+%") )

# Get CM9 baseline data with TPS
no_ne <- filter(data,
                !is.na(PDL1_Av_Day1) &
                !is.na(TPSbaseline) &
                  Dataset == "CM9")

# Count number of datapoints plotted
plotcount <- nrow(no_ne)

#Count the samples
count_0 <- sum(no_ne$TPScat == "0%")
count_1 <- sum(no_ne$TPScat == "1to5%")
count_5 <- sum(no_ne$TPScat == "5+%")

spdl1_vs_TPScat_plot_9 <- ggplot(no_ne,
                          aes(y = PDL1_Av_Day1, x = TPScat)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
            			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 #facet_grid( ~ Dataset) +
  theme_dj(7) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: Tumor Positivity Score Category",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
    				scale_x_discrete(labels=c("0%" =  paste0("0%\nN=",count_0), 
							  "1to5%" =  paste0("1-5%\nN=",count_1),
							  "5+%" =  paste0("5+%\nN=",count_5)))+
  	stat_compare_means(method="wilcox.test", size = 3,
					   aes(label = paste0("P = ", ..p.format..)),
					   comparisons = my_comparisons,
					   vjust = "inward")+
theme(legend.position = "none")

spdl1_vs_TPScat_plot_9
```

## Figure 1 d: sPDL1 versus Timepoint in CM38

```{r Time_paired_boxplot_cm38}

#this tidy data already has null observations removed
no_ne <- tidydata_38

# Visualize: Specify the comparisons I want
my_comparisons <- list( c("Baseline", "Day 29"),
                        c("Baseline", "Day 43"))

#count samples in data
plotcount <- nrow(no_ne)

#Count the samples
count_baseline <- sum(no_ne$Day == "Baseline")
count_day29 <- sum(no_ne$Day == "Day 29")
count_day43 <- sum(no_ne$Day == "Day 43")


spdl1_byDay_38 <- no_ne %>%
  ggplot(aes(x = Day, y = sPDL1)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQflag, shape = ULOQflag),
            			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 #facet_grid( ~ Dataset) +
  theme_dj(7) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Time",
       y = "sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
      	scale_x_discrete(labels=c("Baseline" = paste0("Baseline\nN=",count_baseline), 
							  "Day 29" = paste0("Day 29\nN=",count_day29),
							  "Day 43" = paste0("Day 43\nN=",count_day63)))+
  	stat_compare_means(method="wilcox.test", size = 3,
					   aes(label = paste0("P = ", ..p.format..)),
					   comparisons = my_comparisons,
					   vjust = "inward")+
theme(legend.position = "none")
  
spdl1_byDay_38
```

## Figure 1 e: Baseline sPDL1 versus Prior Therapy CM38

```{r wilcoxon_day1_PriorTherapyGroup_CM38}

print(wilcox.test(PDL1_Av_Day1 ~ PriorTherapyGroup, data = data38))
```

CM38: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the two PriorTherapyGroup categories are identical populations.

```{r boxplot_baseline_prior_38}

# Specify desired comparison
my_comparisons <- list( c("NIV3-NAIVE", "NIV3-PROG"))

# Count number of datapoints plotted
plotcount <- nrow(data38)

#Count the samples
count_naive <- sum(data38$PriorTherapyGroup == "NIV3-NAIVE")
count_pretreated <- sum(data38$PriorTherapyGroup == "NIV3-PROG")

spdl1_vs_prior_plot_38 <- ggplot(data38,
                          aes(y = PDL1_Av_Day1, x = PriorTherapyGroup)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Prior ipilimumab status",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("NIV3-NAIVE" =  paste0("Naive\nN=",count_naive), 
							  "NIV3-PROG" =  paste0("Experienced\nN=",count_pretreated)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_prior_plot_38

```
## Figure 1 f: Baseline sPDL1 versus Tumor PDL1 Score in CM38

```{r spdl1_vs_TPScat_plot_38}

# Specify desired comparison
my_comparisons <- list( c("0%", "1to5%"), 
 #                       c("1to5%", "5+%"), 
                        c("0%", "5+%") )

# Get CM38 baseline data with TPS
no_ne <- filter(data,
                !is.na(PDL1_Av_Day1) &
                !is.na(TPSbaseline) &
                  Dataset == "CM38")

# Count number of datapoints plotted
plotcount <- nrow(no_ne)

#Count the samples
count_0 <- sum(no_ne$TPScat == "0%")
count_1 <- sum(no_ne$TPScat == "1to5%")
count_5 <- sum(no_ne$TPScat == "5+%")

spdl1_vs_TPScat_plot_38 <- ggplot(no_ne,
                          aes(y = PDL1_Av_Day1, x = TPScat)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
            			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 #facet_grid( ~ Dataset) +
  theme_dj(7) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Tumor Positivity Score Category",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
    				scale_x_discrete(labels=c("0%" =  paste0("0%\nN=",count_0), 
							  "1to5%" =  paste0("1-5%\nN=",count_1),
							  "5+%" =  paste0("5+%\nN=",count_5)))+
  	stat_compare_means(method="wilcox.test", size = 3,
					   aes(label = paste0("P = ", ..p.format..)),
					   comparisons = my_comparisons,
					   vjust = "inward")+
theme(legend.position = "none")

spdl1_vs_TPScat_plot_38
```


\newpage
# Supplemental Figure Results

##CM9 Baseline sPDL1 versus Age

```{r kruskal_day1_age_CM9}

# Get CM9 baseline data
data9bl <- filter(data9,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ age_class, data = data9bl))
```

CM9: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the  two age categories are identical populations.

```{r boxplot_baseline_age_9}


# Specify desired comparison
my_comparisons <- list( c("< 65",  ">= 65"))


# Count number of datapoints plotted
plotcount <- nrow(data9bl)

#Count the samples
count_young <- sum(data9bl$age_class == "< 65")
count_old <- sum(data9bl$age_class == ">= 65")


spdl1_vs_age_plot_9 <- ggplot(data9bl,
                          aes(y = PDL1_Av_Day1, x = age_class)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: Age group",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("< 65" =  paste0("< 65y\nN=",count_young), 
							  ">= 65" =  paste0(">= 65y\nN=",count_old)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 


spdl1_vs_age_plot_9

  

```

##CM9 Baseline sPDL1 versus Sex

```{r kruskal_day1_sex_CM9}

# Get CM9 baseline data
data9bl <- filter(data9,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ SEX, data = data9bl))
```

CM9: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the  two SEX categories are identical populations.

```{r boxplot_baseline_sex_9}

# Specify desired comparison
my_comparisons <- list( c("1",  "2"))

# Count number of datapoints plotted
plotcount <- nrow(data9bl)

#Count the samples
count_m <- sum(data9bl$SEX == "1")
count_f <- sum(data9bl$SEX == "2")

spdl1_vs_sex_plot_9 <- ggplot(data9bl,
                          aes(y = PDL1_Av_Day1, x = SEX)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: Sex",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("1" =  paste0("Male\nN=",count_m), 
							  "2" =  paste0("Female\nN=",count_f)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_sex_plot_9

```


##CM9 Baseline sPDL1 versus PBRM1

```{r kruskal_day1_pbrm1_CM9}

# Get CM9 baseline data
data9bl <- filter(data9,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ PBRM1_class, data = data9bl))
```

CM9: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the three PBRM1 categories are identical populations.

```{r boxplot_baseline_pbrm_9}

# Specify desired comparison
#levels = c("Mutant", "WT", "ND")
my_comparisons <- list( c("Mutant", "WT"))

# Count number of datapoints plotted
plotcount <- nrow(data9bl)

#Count the samples
count_wt <- sum(data9bl$PBRM1_class == "WT")
count_mut <- sum(data9bl$PBRM1_class == "Mutant")
count_nd <- sum(data9bl$PBRM1_class == "ND")

spdl1_vs_PBRM1_plot_9 <- ggplot(data9bl,
                          aes(y = PDL1_Av_Day1, x = PBRM1_class)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: PBRM1 status",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("Mutant" =  paste0("MUT\nN=",count_mut), 
							  "WT" =  paste0("WT\nN=",count_wt),
							  "ND" =  paste0("ND\nN=",count_nd)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_PBRM1_plot_9

```

##CM9 Baseline sPDL1 versus VHL

```{r kruskal_day1_vhl_CM9}

# Get CM9 baseline data
data9bl <- filter(data9,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ VHL_class, data = data9bl))
```

CM9: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the three VHL_class categories are identical populations.

```{r boxplot_baseline_vhl_9}

# Specify desired comparison
#levels = c("Mutant", "WT", "ND")
my_comparisons <- list( c("Mutant", "WT"))

# Count number of datapoints plotted
plotcount <- nrow(data9bl)

#Count the samples
count_wt <- sum(data9bl$VHL_class == "WT")
count_mut <- sum(data9bl$VHL_class == "Mutant")
count_nd <- sum(data9bl$VHL_class == "ND")

spdl1_vs_VHL_plot_9 <- ggplot(data9bl,
                          aes(y = PDL1_Av_Day1, x = VHL_class)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM9: VHL1 status",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("Mutant" =  paste0("MUT\nN=",count_mut), 
							  "WT" =  paste0("WT\nN=",count_wt),
							  "ND" =  paste0("ND\nN=",count_nd)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_VHL_plot_9

```

##CM38 Baseline sPDL1 versus Age

```{r kruskal_day1_age_CM38}

# Get CM38 baseline data
data38bl <- filter(data38,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ AGEGR2, data = data38bl))
```

CM38: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the  two age categories are identical populations.

```{r boxplot_baseline_age_38}


# Specify desired comparison
my_comparisons <- list( c("< 65",  ">= 65"))


# Count number of datapoints plotted
plotcount <- nrow(data38bl)

#Count the samples
count_young <- sum(data38bl$AGEGR2 == "< 65")
count_old <- sum(data38bl$AGEGR2 == ">= 65")


spdl1_vs_age_plot_38 <- ggplot(data38bl,
                          aes(y = PDL1_Av_Day1, x = AGEGR2)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Age group",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("< 65" =  paste0("< 65y\nN=",count_young), 
							  ">= 65" =  paste0(">= 65y\nN=",count_old)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 


spdl1_vs_age_plot_38

  

```

##CM38 Baseline sPDL1 versus Sex

```{r kruskal_day1_sex_CM38}

# Get CM38 baseline data
data38bl <- filter(data38,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ SEX, data = data38bl))
```

CM38: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the  two SEX categories are identical populations.

```{r boxplot_baseline_sex_38}

# Specify desired comparison
my_comparisons <- list( c("M",  "F"))

# Count number of datapoints plotted
plotcount <- nrow(data38bl)

#Count the samples
count_m <- sum(data38bl$SEX == "M")
count_f <- sum(data38bl$SEX == "F")

spdl1_vs_sex_plot_38 <- ggplot(data38bl,
                          aes(y = PDL1_Av_Day1, x = SEX)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Sex",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("M" =  paste0("Male\nN=",count_m), 
							  "F" =  paste0("Female\nN=",count_f)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_sex_plot_38
```

##CM38 Baseline sPDL1 versus Subtype

```{r kruskal_day1_subtype_CM38}

# Get CM38 baseline data
data38bl <- filter(data38,
                !is.na(PDL1_Av_Day1))
  
print(kruskal.test(PDL1_Av_Day1 ~ MELSUBT, data = data38bl))
```

CM38: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the subtype categories are identical populations.

```{r boxplot_baseline_subtype_38}

data38bl$MELSUBT <- factor(data38bl$MELSUBT,
                           levels = c("CUTANEOUS",
                                 "MUCOSAL",
                                 "OCULAR/UVEAL",
                                 "ACRAL",
                                 "OTHER"))

# Specify desired comparison
my_comparisons <- list( c("CUTANEOUS", "MUCOSAL"))

# Count number of datapoints plotted
plotcount <- nrow(data38bl)

#Count the samples
count_cut <- sum(data38bl$MELSUBT == "CUTANEOUS")
count_muc <- sum(data38bl$MELSUBT == "MUCOSAL")
count_uv <- sum(data38bl$MELSUBT == "OCULAR/UVEAL")
count_ac <- sum(data38bl$MELSUBT == "ACRAL")
count_other <- sum(data38bl$MELSUBT == "OTHER")

spdl1_vs_subt_plot_38 <- ggplot(data38bl,
                          aes(y = PDL1_Av_Day1, x = MELSUBT)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: Subtype",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("CUTANEOUS" =  paste0("Cut\nN=",count_cut), 
							  "MUCOSAL" =  paste0("Muc\nN=",count_muc),
							  "OCULAR/UVEAL" =  paste0("Uveal\nN=",count_uv),
							  "ACRAL" =  paste0("Acral\nN=",count_ac),
							  "OTHER" =  paste0("Other\nN=",count_other)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_subt_plot_38
```

##CM38 Baseline sPDL1 versus LDH

```{r kruskal_day1_LDH_CM38}

# Get CM38 baseline data
data38bl_ldh <- filter(data38,
                !is.na(PDL1_Av_Day1),
                LDHCAT1 != "NOT REPORTED")
  
print(kruskal.test(PDL1_Av_Day1 ~ LDHCAT1, data = data38bl_ldh))
print(kruskal.test(PDL1_Av_Day1 ~ LDHCAT2, data = data38bl_ldh))

```

CM38: At .05 significance level, I conclude that the PDL1_Av_Day1 levels in the LDHCAT1 categories are NOT identical populations.

```{r boxplot_baseline_ldh_38}

data38bl_ldh$LDHCAT1 <- factor(data38bl_ldh$LDHCAT1,
                           levels = c("<= ULN",
                                      "> ULN"))

# Specify desired comparison
my_comparisons <- list( c("<= ULN","> ULN"))

# Count number of datapoints plotted
plotcount <- nrow(data38bl_ldh)

#Count the samples
count_1 <- sum(data38bl_ldh$LDHCAT1 == "<= ULN")
count_2 <- sum(data38bl_ldh$LDHCAT1 == "> ULN")

spdl1_vs_ldh_plot_38 <- ggplot(data38bl_ldh,
                          aes(y = PDL1_Av_Day1, x = LDHCAT1)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(aes(color = ULOQ_FLAG_Day1, shape = ULOQ_FLAG_Day1),
             			   size = 0.5,stroke = 1, 
             position = position_jitter(width = 0.3, height = 0)) +
 scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
scale_color_manual(values = flagColors) +
   scale_shape_manual(values = flagshape_empty) +
  labs(x = "CM38: LDH class",
       y = "Baseline sPDL1, pg/ml",
       shape = "BOR", color = "BOR") +
  				scale_x_discrete(labels=c("<= ULN" =  paste0("< ULN\nN=",count_1), 
							  "> ULN" =  paste0("> ULN\nN=",count_2)))+
    theme_dj(7) +
theme(legend.position = "none")+
 stat_compare_means(comparisons = my_comparisons,
                    method = "wilcox.test",
                    label = "p.format",
                    vjust = "inward",
                    size = 3) 

spdl1_vs_ldh_plot_38
```

\newpage
# Outputs

```{r cowplot_figure}

## Make the Figure pdf
fig1_file <- paste(results_dir, "/Figure1_sPDL1_Clinical_Characteristics.pdf",
  sep="")

pdf(file=fig1_file, width = 7, height = 9)

cowplot::plot_grid(NULL,
                   NULL,
                   NULL,
                   spdl1_byDay_9,
                   spdl1_vs_prior_plot_9,
                  spdl1_vs_TPScat_plot_9,
                  spdl1_byDay_38,
                    spdl1_vs_prior_plot_38,
                  spdl1_vs_TPScat_plot_38,
                    ncol=3,
                   nrow = 3,
                   scale = c(1,1,1,1,1, 1,1,1,1),
                   rel_heights = c(1,1,1),
                   rel_widths = c(1,0.9,1),
                   labels = c("A","","","B","C","D","E","F","G"))

dev.off()

```


```{r cowplot_supplemental}
## Make the Supplemental pdf
fig1_file <- paste(results_dir, "/FigureS1_sPDL1_more_Clinical_Characteristics.pdf",
  sep="")

pdf(file=fig1_file, width = 7, height = 9)

cowplot::plot_grid(spdl1_vs_age_plot_9,
                   spdl1_vs_sex_plot_9,
                   spdl1_vs_PBRM1_plot_9,
                   spdl1_vs_VHL_plot_9,
                   spdl1_vs_age_plot_38,
                   spdl1_vs_sex_plot_38,
                   spdl1_vs_subt_plot_38,
                   spdl1_vs_ldh_plot_38,
                    ncol=4,
                   nrow = 3,
                   scale = c(1,1, 1,1,1,1),
                   rel_heights = c(1,1,1),
                   rel_widths = c(1,1,1.1),
                   labels = c("a","b","c","d", "e","f","g", "h"))

dev.off()

```

