---
title: "Figure 4: sPDL1  versus Affymetrix baseline in Melanoma"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clear the environment
rm(list = ls())
# Free up memory by forcing garbage collection
invisible(gc())  
# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 38
##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

# Paths, Packages, Functions, Palettes

```{r Paths}

# Provide paths
data_dir <- "./Data"
results_dir <- "./Outputs/Manuscript"

```


```{r LoadPackages}
library(knitr) # report generation
library(ggpubr)

library(RColorBrewer)
library(limma)

library(ComplexHeatmap)
library(circlize)

library(grid)
library(gtable)

library(cowplot)

##_ alternatively load tidyverse which bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
##_ use suppressPackageStartup() to hide package load messages
suppressPackageStartupMessages(library(tidyverse))  
```

```{r functions}

source("./Code/Code_PostBMS/theme_dj.R")

```

```{r Palettes}

# color map for response
borColors <- c("PD" = "darkred", "SD" = "darkblue", "CRPR" = "darkgreen", "NE" = "grey")

# color map for response: DFCI colors
# ORANGE, AQUA, CORNFLOWER BLUE, GREY
borColors_DFCI <- c("PD" = "#ff9900", 
               "SD" = "#4a86e8", 
               "CRPR" = "#00ffff",
               "NE" = "#999999")

# Colorblind palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Color for GSEA enrichment
NES_direction <- c("Positive" = "firebrick",
                   "Negative" = "dodgerblue")


# shape map for response, use stroke = 2 to thicken line
borshape_empty <- c("PD" = 0, 
               "SD" = 1, 
               "CRPR" = 2,
               "NE" = 10)

# shape map for response
borshape_solid <- c("PD" = 15, 
               "SD" = 16, 
               "CRPR" = 17,
               "NE" = 10)
```

\newpage
# Objective

1. baseline sPDL1 vs Signatures
1. baseline sPDL1 vs Affy GSEA
1. baseline sPDL1 vs Leading edge, Marco?


## Resources


## Data Sources

```{r loaddata}

# Expression values
rma_file <- paste(data_dir, "CA209038-tumorAffy-HGU219_HS_ENTREZG.rma", sep = "/")
rma <- read_tsv(rma_file)

# Make rma Colnames match "Assay.Name" in sdrf annotation
colnames(rma) <- sub("ENTREZG-", "", colnames(rma))
rma <- rename(rma, Probeset = X1)


# SDRF (Sample and Data Relationship Format) file from Array Express with FA01 update
#FA01 = final analysis lock 01
#CM0038_annotation_of_Affymetrix_sPDL1.rmd
sdrf_file_38 <- paste(data_dir, "submission2439_annotare_2019sdrf_sPDL1_Sigs_CS.Rds", sep = "/")
sdrf_38 <- readRDS(sdrf_file_38)



# # Affymetrix probeset to Gene annotation
probeset_file <- paste(data_dir, "probeset_annotation.rds", sep = "/")
probeset <- readRDS(probeset_file)

#GSEA_Hallmark results for both trials
GSEA_file <- paste(data_dir, "GSEA_Hallmark_CM9_CM38.txt", sep = "/")
GSEA <- read_tsv(GSEA_file)


#MMP results for both trials
MMP_file <- paste(data_dir, "MMPs_CM9_CM38_BLlimmaResults.txt", sep = "/")
MMP <- read_tsv(MMP_file)
```

Affymetrix RMA data was loaded from:

+ *`r rma_file`*

Affymetrix sample annotation tied to USUBJID (Sample and Data Relationship Format, SDRF file) was loaded from:

+ *`r sdrf_file`*

Affymetrix Probeset Annotation was loaded from:

+ *`r probeset_file`*

GSEA results against Hallmark pathways from Long Yuan, merged data 

+ *`r GSEA_file`*

Baseline sPDL1 limma results against MMP list from Kathleen Mahoney, merged data 

+ *`r MMP_file`*

```{r median_values}

bl.median_38 <- 2312

```

Median baseline value calculated on the entire dataset (eg Table 1) was stored for barplots.

+ bl.median_38 <- 2312



## SDRF Screen

```{r screen_rma_datasets}

# Baseline Subject sample 
sdrfScreenSpdl1 <- sdrf_38 %>%
	filter(sampling_time_point == "screen" &
	         !is.na(PDL1_Av_Day1))

# Select 44 Screen Array columns plus Probeset column
# 'Select' gets Assay_Name cols in the same order as the sdrf annotation
# Order is Vital for ComplexHeatmap!! Vital for limma!

rmascreen <- select(rma, one_of("Probeset",sdrfScreenSpdl1$Assay_Name))

#remove total rma data to free up memory
rm(rma)

```


\newpage
# Results


## CM38 Correlation: baseline sPDL1 to baseline numeric scores

For CIBERSORT, I am using pre-treatment columns where more than 20% of samples have a non-zero value.

```{r CM38_correlation_baseline_baseline}

#here are all the pretreatment cibersort columns
#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1%>%
  select(starts_with("CS_pre"))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c("IM150_Angio.Score",
		   "IM150_MyeloidInfl.Score",
		   "TIS",
	"EMTstroma.Score",
	"Cytolytic.Score",
	cibersort_columns)


cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#complete sPDL1
ScreenSpdl1 <- sdrfScreenSpdl1%>%
  filter(!is.na(PDL1_Av_Day1))

#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

correlation <- cor.test(ScreenSpdl1$PDL1_log2_Day1, y = ScreenSpdl1[[i]],
                    method = j)

p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))

}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}
  
kable(arrange(df_data, p.value_pearson),
      title = "CM38: Correlation for Baseline sPDL1 versus Affymetrix scores")

```


```{r scatterplot_significant}

significant_scores <- as.character(df_data$Biomarker[df_data$p.value < .05])
  
for(i in significant_scores){
  
  scatterplot <- ScreenSpdl1%>%
    ggplot(aes(y = PDL1_Av_Day1, x = .data[[i]]))+
	geom_point(aes(y = PDL1_Av_Day1, x = .data[[i]], 
				   colour = BOR3, shape = BOR3),
			   size = 1,stroke = 1,
			   position = position_jitter(0)) +
	stat_smooth(method = "lm", se=TRUE, color="black", formula = y ~ x) +
    scale_color_manual(values = borColors_DFCI) +
  scale_shape_manual(values = borshape_empty)+
  theme_dj(8) +
   scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
  labs(y = "Baseline sPDL1, pg/ml",
       x = paste("Baseline score:", i)) +
  # geom_text(data=data.frame(x=4,y=1024),
  # 		  aes(x, y),
  # 		  label= paste("Adj P =",signif(p_value, 2)),
  # 		  size = 3,
  # 		  hjust = "inward")+
  theme(legend.position = "none")
  
print(scatterplot)  
}

```

Barplot shows baseline sPDL1 colored by IM150_Angio.

```{r CM38_barplot_baseline_Angio}


plotcount <- nrow(sdrfScreenSpdl1)

correlation <- cor.test(sdrfScreenSpdl1$PDL1_log2_Day1, 
                        y = sdrfScreenSpdl1$IM150_Angio.Score,
                    method = "spearman")
  
 
p.value <- round(correlation$p.value[1], 3)
estimate <- round(correlation$estimate[[1]], 2)

#Find number of patients above median value on all patients
median_patient <- sum(ScreenSpdl1$PDL1_Av_Day1 > bl.median_38)

barplot_baseline_angio_38 <- sdrfScreenSpdl1 %>%
			ggplot(aes(x = as.factor(reorder(individual, desc(PDL1_Av_Day1))), 
			           y = PDL1_Av_Day1, 
			           fill = IM150_Angio.Score)) +
geom_bar(stat="identity",
			           color = "black",
         size = 0.25) +
  # facet_grid(~ccrccCluster, 
  #             scales = 'free',
  #              space = "free")+
  scale_fill_gradient2(
  low = "navyblue",
  mid = "white",
  high = "firebrick",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill") +
  coord_cartesian(ylim=c(512,16384))+
 		geom_vline(xintercept= median_patient, linetype="dashed", 
                color = "grey", size=1)+
   scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    # limits = c(512,16384),
                    n.breaks = 6) +
  labs(#subtitle = paste("Patients with Screen Affymetrix and baseline sPDL1, N=",plotcount),
		 x = "CM038 Patient, colored by Angio Score",
		 y = "Baseline sPDL1, pg/ml",
		 fill = "IM150 Angio Score") +
      annotate("text", x=33, y=8192, 
           label= paste0("Spearman p = ",p.value, 
                         "\nR = ", estimate), 
          color = "black",
           size = 3)+
        annotate("text", x=median_patient, y=16384, 
           label= "Above  Below", 
           color = "black",
           size = 3)+
  theme_dj(8)+
  theme(legend.position = "bottom",
		  axis.text.x=element_text(size=6, angle = 90, vjust = 0.5, hjust=1),
		  axis.title=element_text(size=8,face="bold")) 

print(barplot_baseline_angio_38)
```


Barplot shows baseline sPDL1 colored by IM150_MyeloidInfl.Score

```{r CM38_barplot_baseline_myeloid}


plotcount <- nrow(sdrfScreenSpdl1)

correlation <- cor.test(sdrfScreenSpdl1$PDL1_log2_Day1, 
                        y = sdrfScreenSpdl1$IM150_MyeloidInfl.Score,
                    method = "spearman")
  
 
p.value <- round(correlation$p.value[1], 3)
estimate <- round(correlation$estimate[[1]], 2)

#Find number of patients above median value on all patients
median_patient <- sum(ScreenSpdl1$PDL1_Av_Day1 > bl.median_38)

barplot_baseline_myeloid_38 <- sdrfScreenSpdl1 %>%
			ggplot(aes(x = as.factor(reorder(individual, desc(PDL1_Av_Day1))), 
			           y = PDL1_Av_Day1, 
			           fill = IM150_MyeloidInfl.Score)) +
geom_bar(stat="identity",
			           color = "black",
         size = 0.25) +
  # facet_grid(~ccrccCluster, 
  #             scales = 'free',
  #              space = "free")+
  scale_fill_gradient2(
  low = "navyblue",
  mid = "white",
  high = "firebrick",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill") +
  coord_cartesian(ylim=c(512,16384))+
 		geom_vline(xintercept= median_patient, linetype="dashed", 
                color = "grey", size=1)+
   scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    # limits = c(512,16384),
                    n.breaks = 6) +
  labs(#subtitle = paste("Patients with Screen Affymetrix and baseline sPDL1, N=",plotcount),
		 x = "CM038 Patient, colored by Baseline IM150 Myeloid Score",
		 y = "Baseline sPDL1, pg/ml",
		 fill = "IM150 Myeloid Score") +
      annotate("text", x=33, y=8192, 
           label= paste0("Spearman p = ",p.value, 
                         "\nR = ", estimate), 
          color = "black",
           size = 3)+
        annotate("text", x=median_patient, y=16384, 
           label= "Above  Below", 
           color = "black",
           size = 3)+
  theme_dj(8)+
  theme(legend.position = "bottom",
		  axis.text.x=element_text(size=6, angle = 90, vjust = 0.5, hjust=1),
		  axis.title=element_text(size=8,face="bold")) 

print(barplot_baseline_myeloid_38)
```

Barplot shows baseline sPDL1 colored by CS_pre.Neutrophil

```{r CM38_barplot_baseline_CS_pre.Neutrophil}


plotcount <- nrow(sdrfScreenSpdl1)

correlation <- cor.test(sdrfScreenSpdl1$PDL1_log2_Day1, 
                        y = sdrfScreenSpdl1$CS_pre.Neutrophil,
                    method = "spearman")
  
 
p.value <- round(correlation$p.value[1], 3)
estimate <- round(correlation$estimate[[1]], 2)

#Find number of patients above median value on all patients
median_patient <- sum(ScreenSpdl1$PDL1_Av_Day1 > bl.median_38)

barplot_baseline_neut_38 <- sdrfScreenSpdl1 %>%
			ggplot(aes(x = as.factor(reorder(individual, desc(PDL1_Av_Day1))), 
			           y = PDL1_Av_Day1, 
			           fill = CS_pre.Neutrophil)) +
geom_bar(stat="identity",
			           color = "black",
         size = 0.25) +
  # facet_grid(~ccrccCluster, 
  #             scales = 'free',
  #              space = "free")+
 scale_fill_gradient2(
  low = "white",
  mid = "white",
  high = "black",
  midpoint = .001,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill") +
  coord_cartesian(ylim=c(512,16384))+
 		geom_vline(xintercept= median_patient, linetype="dashed", 
                color = "grey", size=1)+
   scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    # limits = c(512,16384),
                    n.breaks = 6) +
  labs(#subtitle = paste("Patients with Screen Affymetrix and baseline sPDL1, N=",plotcount),
		 x = "CM038 Patient, colored by Baseline Neutrophil level",
		 y = "Baseline sPDL1, pg/ml",
		 fill = "CIBERSORT Neutrophil") +
      annotate("text", x=33, y=8192, 
           label= paste0("Spearman p = ",p.value, 
                         "\nR = ", estimate), 
          color = "black",
           size = 3)+
        annotate("text", x=median_patient, y=16384, 
           label= "Above  Below", 
           color = "black",
           size = 3)+
  theme_dj(8)+
  theme(legend.position = "bottom",
		  axis.text.x=element_text(size=6, angle = 90, vjust = 0.5, hjust=1),
		  axis.title=element_text(size=8,face="bold")) 

print(barplot_baseline_neut_38)
```



## A Heatmap: baseline sPDL1, Signatures, Cibersort

```{r heatmap_sPDL1_sigs_Cibersort}

#sort data by sPDL1
#make a new object or it messes up all future complexheatmaps
sdrfScreenSpdl1sort <- sdrfScreenSpdl1%>%
	arrange(desc(PDL1_Av_Day1))

############# Signatures heatmap matrix
mat <- sdrfScreenSpdl1sort%>%
	select("EMTstroma.Score",
		   "TIS",
    "Cytolytic.Score",
    "IM150_Angio.Score",
    "IM150_MyeloidInfl.Score")

# transpose the matrix
mat <- t(mat)

# Edit the rownames
row.names(mat) <- gsub( ".Score", "", row.names(mat))
row.names(mat) <- gsub( "IM150_", "", row.names(mat))
row.names(mat) <- gsub( "Angio", "Angiogenesis", row.names(mat))
row.names(mat) <- gsub( "Infl", "Inflammation", row.names(mat))

# modify rowname for significant correlation
row.names(mat)  <- gsub( "Inflammation", "Inflammation (p = 0.046)", row.names(mat))


########## cibersort heatmap matrix

#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1sort%>%
  select(starts_with("CS_pre."))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#if you only want to use columns where more than e.g.20% of samples have a value, select them
mat_cs <- select(mat_cs,
       one_of(cibersort_columns))


#Scale each column as a fraction of the highest value (100)
#https://stackoverflow.com/questions/19462321
mat_cs_scale <- data.frame(lapply(mat_cs, function(x) scale(x, center = FALSE, scale = max(x, na.rm = TRUE)/100)))


# transpose the matrix and modify rownames
mat_cs_scale <- t(mat_cs_scale)
row.names(mat_cs_scale) <- gsub( "CS_pre.", "", row.names(mat_cs_scale))
row.names(mat_cs_scale) <- gsub( "\\.", "_", row.names(mat_cs_scale))
row.names(mat_cs_scale) <- gsub( "dendritic_cell", "DC", row.names(mat_cs_scale))



########### Heatmap components
#conditional coloring in barplot removed
						   						# gp = gpar(fill = 
						   						# 		  	ifelse(sdrfScreenSpdl1sort$BOR3 == "PD",
						   						# 		  		   "#ff9900", "black"))),


# Want to plot BOR3, sPDL1 
ha_top = HeatmapAnnotation(sPDL1 = anno_barplot(sdrfScreenSpdl1sort$PDL1_Av_Day1, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill ="black")),
						   blank = sdrfScreenSpdl1sort$sampling_time_point,
						   response = sdrfScreenSpdl1sort$BOR3,
					   col = list(
						     response = c("PD" = "#ff9900", 
               "SD" = "#4a86e8", 
               "CRPR" = "#00ffff"),
							  blank = c("screen" = "white")),
						   annotation_height = unit(c(2,0.1, 0.3), "cm"),
					   						   show_annotation_name = c(sPDL1 = TRUE,
					   						                            blank = FALSE,
					   						                            response = TRUE),
						 show_legend = FALSE)


#draw(ha_top,1:56)
#draw(ha_bottom,1:56)

heatmap_object1 = Heatmap(mat,
						 col = colorRamp2(c(-1, 0, 1), 
						 				 c("dodgerblue", "white", "firebrick")), 
						 cluster_rows = TRUE,
						 clustering_distance_rows = "euclidean",
    					clustering_method_rows = "ward.D2",
						    row_dend_side = "left",
    				row_dend_width = unit(5, "mm"),
    				show_row_dend = TRUE,
						 km = 2,
						 show_row_names = TRUE,
						 row_names_side = "right",
						 row_names_gp = gpar(fontsize = 8),
						 cluster_columns = FALSE,
#						 column_title = "CM38: 44 Baseline, Publication Signature Correlation",
						 show_column_names = FALSE,
						 height = unit(1.5, "cm"),
						 width = unit(12, "cm"),
						show_heatmap_legend = FALSE)

heatmap_object2 = Heatmap(mat_cs_scale,
					col = colorRamp2(c( 0, 100), 
						  				 c( "white", "black")), 
						 cluster_rows = TRUE,
						 clustering_distance_rows = "euclidean",
    					clustering_method_rows = "ward.D2",
						    row_dend_side = "left",
    				row_dend_width = unit(5, "mm"),
    				show_row_dend = TRUE,
						 km = 2,
						 show_row_names = TRUE,
						 row_names_side = "right",
						 row_names_gp = gpar(fontsize = 8),
						 cluster_columns = FALSE,
#						 column_title = "CM38: 44 Baseline, Publication Signature Correlation",
						 show_column_names = FALSE,
						 height = unit(4.5, "cm"),
						 width = unit(12, "cm"),
						show_heatmap_legend = FALSE)

ht_list = ha_top %v% heatmap_object1 %v% heatmap_object2 

draw(ht_list)

# Decorate the heatmap object, adding layers of texts or lines
decorate_annotation("sPDL1", 
					{grid.text("Baseline sPDL1, pg/ml",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 6))})

# https://github.com/jokergoo/ComplexHeatmap/issues/110

gb_heatmap_cs = grid.grabExpr(draw(ht_list,
                                padding = unit(c(2, 20, 2, 30), "mm")),
                              decorate_annotation("sPDL1", 
					{grid.text("Baseline sPDL1, pg/ml",
							   unit(-10, "mm"), just = "bottom",
							   rot = 90, check.overlap = T,
							   gp = gpar(fontsize = 6))}))


#To use cowplot, convert the heatmap gtree returned from grid.grabExpr into a gtable:

heatmap_gtable_cs <- gtable_matrix("heatmap_gtable", matrix(list(gb_heatmap_cs)), unit(1, "null"), unit(1, "null"))

```


\newpage
## limma vs baseline sPDL1

Run limma on 44 SCREEN samples, with baseline sPDL1 (log2) as continuous variable

```{r limma_baseline_spdl1}

# make Probeset column into rownames
data <- rmascreen %>% remove_rownames %>% column_to_rownames(var="Probeset")

# Run limma
design <- model.matrix(~PDL1_log2_Day1,sdrfScreenSpdl1) 
fit <-lmFit(data, design)
fit <- eBayes(fit)

kable(summary(decideTests(fit)),
	  title = "CM38 baseline: Summary of regression against PDL1_log2_Day1")

```

By a Benjamini-Hochberg false discovery rate (FDR) there are no results that pass multiple test correction.

```{r annotate_limma_results}
# Annotate association hit results
results <- topTable(fit, coef=2,  number=Inf, sort.by = "B") %>% 
	rownames_to_column(var="Probeset") %>%
	left_join(dplyr::select(probeset, -matches("Geneset")))%>%
				arrange(desc(logFC)) 

```


```{r iris_spdl1}


# Make a summary by direction of correlation and IRIS classes
upsum <- results %>%
	filter(logFC >0) %>%
	count(IRIS_Most_Specific) %>%
	dplyr::rename("PositiveAssociation w/ sPDL1" = n)

downsum  <- results %>%
	filter(logFC < 0) %>%
	count(IRIS_Most_Specific) %>%
	dplyr::rename("NegativeAssociation w/ sPDL1"  = n)

irisassociation_spdl <- left_join(upsum,downsum) %>%
	arrange(match(IRIS_Most_Specific,c("T Cell","B Cell","NK Cell","Lymphoid","Dendritic Cell","Monocyte","Neutrophil","Myeloid","Multiple",NA)))%>%
	dplyr::rename("Immune annotation (IRIS)"  = "IRIS_Most_Specific")
	
kable(irisassociation_spdl,
	  title = "CM38 Baseline: Immune class Direction of association with sPDL1")
```


## Bonferroni Pvalue correction

However since many transcripts are coregulated or reflect cell types, the assumption of independence is invalid. Instead we would like to correct for the true number of independent statistical comparisons. Following previous studies, this can be estimated by identifying the number of principal components accounting for 95% of the variance in the data (Gao et al., 2008; Johnson et al., 2010; Tromp et al., 2018). We apply this PC-based Bonferroni correction to derive the level of significance.


```{r cumulative_variance}

# PC-based multiple testing correction method
pc <- prcomp(t(data), scale.=FALSE)  # RMA is already log2 scaled?

# Calculate the proportion of variance accounted for by each PC
proportionVariance <- (pc$sdev ^ 2) / sum(pc$sdev ^ 2)

# Find number of PCs with cumulative proportion of variance of at least 95%
nPC95Var <- min(which(cumsum(proportionVariance) > 0.95))

rm(pc, proportionVariance)  # clear namespace

```

The number of PCs with cumulative proportion of variance of at least 95% is *`r nPC95Var`* (37). We use this estimate as the number of independent statistical comparisons for Bonferroni correction to derive the level of significance.


```{r p_adj_pc_bonferroni}


#Do math here to get PC-based Bonferroni correction 

results<- results%>%
	mutate(p_adj_pc_bonferroni = P.Value*nPC95Var)

#fix P>1
results$p_adj_pc_bonferroni <-	if_else(results$p_adj_pc_bonferroni>1, 1,results$p_adj_pc_bonferroni)


#compute gsea_rank for Ron's GSEA
results <- results%>%
  mutate(gsea_rank=(sign(logFC) * -log10(P.Value)))

```

I extracted and annotated the results table for all 18571 probesets in the sPDL1 limma results.

There are *`r sum(results$p_adj_pc_bonferroni < 0.05)`* (9) genes which pass P<0.05 in a PCA-Bonferroni adjustment.

\newpage

Results by Bonerroni adjP <0.05 and Fold > 1.2 change in expression level per unit increase of sPDL1 value.(sPDL1 value is log2 so 1 unit is 2-fold) = 2 genes.

\footnotesize

```{r results_baseline_05}
#p_adj_pc_bonferroni <0.05 
#abs(logFC) > 0.2630344 means Fold > 1.2 change in expression level per unit increase of sPDL1 value
#this is a stringent fold since there are about 4 units in the log2 scale
resultsp05 <- results %>% 
					   filter(p_adj_pc_bonferroni < 0.05,
					        abs(logFC) > 0.2630344) %>%
				arrange(desc(logFC)) 


kable(resultsp05,
      caption = "CM38 limma on Baseline sPDL1: Bonf_P<0.05 and Fold > 1.2 per 2-fold sPDL1")

```

\normalsize

\newpage
## GDF15 association with Baseline sPDL1

Macrophage inhibitory cytokine-1 (MIC-1/GDF15), a divergent member of the TGF-β superfamily, is over-expressed by many common cancers including those of the prostate (PCa) and its expression is linked to cancer outcome. 

```{r scatterplot_baseline_positive, fig.height= 5, fig.width=5 }


#"51083_at" GAL = galanin and GMAP prepropeptide
# 9518_at gdf15
# 6422_at = sfrp1
# 3576_at = IL8

# Define the desired probeset and grab the gene Symbol

probesetwanted <- "9518_at"
genewanted <- as.character(probeset[probeset$Probeset == probesetwanted, 4])

# Get Gene values, then transpose to Array ID and RMA values in columns
onegene <- rmascreen %>% 
	filter(rmascreen$Probeset == probesetwanted) %>%
	gather(key = "Assay_Name", value = "RMA", -Probeset) %>%
	left_join(select(sdrfScreenSpdl1, 
					 c("Assay_Name", "OR_group","BOR3", 
					   "individual", "PDL1_Av_Day1")),
					 )

#Get P value
p_value <- results[results$Probeset == probesetwanted, 12 ]

#plot     
scatterplot_limma <- onegene %>%
	ggplot(aes(y = PDL1_Av_Day1, x = RMA))+
	geom_point(aes(y = PDL1_Av_Day1, x = RMA, 
				   colour = BOR3, shape = BOR3),
			   size = 1,stroke = 1,
			   position = position_jitter(0)) +
	stat_smooth(method = "lm", se=TRUE, color="black", formula = y ~ x) +
    scale_color_manual(values = borColors_DFCI) +
  scale_shape_manual(values = borshape_empty)+
  theme_dj(8) +
   scale_y_continuous(trans = 'log2',
                    breaks = waiver(),
                    n.breaks = 6) +
  labs(y = "Baseline sPDL1, pg/ml",
       x = "GDF15: Baseline expression signal, RMA")+
  		  geom_text(data=data.frame(x=5,y=1024), 
  		  aes(x, y), 
  		  label= paste("Adj P =",signif(p_value, 2)),
  		  size = 3,
  		  hjust = "inward")+
  theme(legend.position = "none")


print(scatterplot_limma)
```

## B: NES plot

```{r barplot_NES_shared}

#https://stackoverflow.com/questions/45843759/ggplot2-coloring-axis-text-on-a-faceted-plot
#it's not possible to color by direction using a theme, with the facets
#one can't map aesthetics into theme(), which controls the appearance of axis text in ggplot.

# New facet label names for Compare_Direction variable
dir.labs <- c("Same as CM9", "Opposite to CM9")
names(dir.labs) <- c("Same", "Opposite")

#filter to genesets shares with CM38
gsea_38 <- GSEA%>%
  filter(Dataset == "CM38",
         !is.na(Dir_in_other_set))%>%
  arrange(NES)

#label_color <- ifelse(gsea_38$Compare_Direction == "Same", "red", "black")

barplot_nes <- gsea_38 %>%
			ggplot(aes(x = as.factor(reorder(Name, NES)), 
			           y = NES, 
			           fill = Dir_in_dataset)) +
    geom_segment( aes(reorder(Name, NES), xend=Name, y=0, yend=NES)) +
  geom_point( size=2, aes( fill = Dir_in_dataset), shape=21, stroke=1) +
       scale_fill_manual(values = NES_direction) +
  coord_cartesian(ylim=c(-2,3))+
  labs(#subtitle = "CM38 Hallmark (shared w CM9)",
		 x = "",
		 y = "Normalized Enrichment Score") +
  theme_dj(8)+
    coord_flip() +  # flip coordinates (puts labels on y axis)
    facet_grid(rows = vars(Compare_Direction),
             scales = "free",
             space = "free_y",
             labeller = labeller(Compare_Direction=dir.labs))+
  theme(legend.position = "none",
		  axis.text.x=element_text(size=8, angle = 0, vjust = 0.5, hjust=1),
		  axis.text.y=element_text(size=8, vjust = 0.5, hjust=1, 
		                           #color = label_color,
		                           face="bold"),
		  axis.title=element_text(size=8,face="bold"))+
	guides(fill=guide_legend(nrow=1,byrow=TRUE)) 
	
print(barplot_nes)

```

```{r barplot_NES_all_hallmark}

#filter to genesets shares with CM38
gsea_38 <- GSEA%>%
  filter(Dataset == "CM38")%>%
  arrange(NES)

label_color <- ifelse(is.na(gsea_38$Dir_in_other_set), "red", "black")

barplot_nes_all <- gsea_38 %>%
			ggplot(aes(x = as.factor(reorder(Name, NES)), 
			           y = NES, 
			           fill = Dir_in_dataset)) +
    geom_segment( aes(reorder(Name, NES), xend=Name, y=0, yend=NES)) +
  geom_point( size=2, aes( fill = Dir_in_dataset), shape=21, stroke=1) +
       scale_fill_manual(values = NES_direction) +
  coord_cartesian(ylim=c(-2,3))+
  labs(subtitle = "CM38 Hallmark (All hits)",
		 x = "",
		 y = "Normalized Enrichment Score") +
  theme_dj(8)+
    coord_flip() +  # flip coordinates (puts labels on y axis)
  theme(legend.position = "none",
		  axis.text.x=element_text(size=8, angle = 0, vjust = 0.5, hjust=1),
		  axis.text.y=element_text(size=8, vjust = 0.5, hjust=1, color = label_color,face="bold"),
		  axis.title=element_text(size=8,face="bold")) 
	guides(fill=guide_legend(nrow=1,byrow=TRUE)) 
	
print(barplot_nes_all)

```
## C: Heatmap Leading edge geneset

```{r identify_genes}

#https://stackoverflow.com/questions/42867124/comma-separated-string-split
#b <- unlist(strsplit(a, ","))

#get the leading edge genes from CM9
genes_9 <- GSEA$core_enrichment[which(GSEA$Name == "IL6_JAK_STAT3_SIGNALING" &
                                        GSEA$Dataset == "CM9")]
genes_9 <-  unlist(strsplit(genes_9, "/"))
  
#get the leading edge genes from CM38
genes_38 <- GSEA$core_enrichment[which(GSEA$Name == "IL6_JAK_STAT3_SIGNALING" &
                                        GSEA$Dataset == "CM38")]
genes_38 <-  unlist(strsplit(genes_38, "/"))

#find overlap
genes_shared <- genes_9[genes_9%in%genes_38]

#get probesets for shared leading edge
probeset_shared <- probeset$Probeset[probeset$Symbol%in%genes_shared]

```


```{r heatmap_leadingEdge}

# # Object of the geneset probesets by significance
# geneset <- filter(results,
#                   logFC < -0.5) %>%
# 	pull(Probeset)

# Subset rmascreen data to the desired geneset 
rma_geneset <- rmascreen[rmascreen$Probeset %in% probeset_shared,]

# Add the Gene Symbol to the geneset expression data table 
# so it can be used to label heatmap rows
rma_geneset <-  probeset %>%
	select(Probeset,Symbol) %>%
	inner_join(rma_geneset)

# scaled_mat will be the body of the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(select(rma_geneset, -Probeset, -Symbol))))
rownames(scaled_mat) = rma_geneset$Symbol

# Provide column order as Ascending sPDL1
# This works because scaled_mat columns are in same order as SDRF rows
ordering = order(sdrfScreenSpdl1$PDL1_Av_Day1,
                 decreasing = TRUE)

# Top heatmap annotation is PDL1_Av_Day1 
ha_top = HeatmapAnnotation(barplot1 = anno_barplot(sdrfScreenSpdl1$PDL1_Av_Day1, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = "black")),
						   annotation_height = unit(c(1), "cm"),
						   show_annotation_name = c(barplot1 = FALSE),
						 show_legend = FALSE)

#draw(ha_top,1:56)
# No Bottom heatmap annotation 

heatmap_object2 = Heatmap(scaled_mat,
						 col = colorRamp2(c(-1, 0, 1), c("dodgerblue", "white", "firebrick")), 
						 cluster_rows = FALSE,
						 km = 1,
						 row_names_gp = gpar(fontsize = 6),
						    				show_row_dend = FALSE,
						 show_row_names = TRUE,
						 row_names_side = "left",
						 cluster_columns = FALSE,
						 column_order = ordering,
						 column_title = "IL6_JAK_STAT3_SIGNALING",
						 column_title_gp = gpar(fontsize = 7),
						 show_column_names = FALSE,
						 width = unit(6, "cm"),
						 						show_heatmap_legend = FALSE,
						 height = unit(3, "cm"),
						 top_annotation = ha_top) 

# https://github.com/jokergoo/ComplexHeatmap/issues/110

gb_heatmap2 = grid.grabExpr(draw(heatmap_object2,
                                padding = unit(c(2, 10, 2, 2), "mm")))

#To use cowplot, convert the heatmap gtree returned from grid.grabExpr into a gtable:

heatmap_gtable2 <- gtable_matrix("heatmap_gtable", matrix(list(gb_heatmap2)), unit(1, "null"), unit(1, "null"))

```


## D: Heatmap MMP genesets

I'm going to plot the genes where t>1.5 in the limma against baseline sPDL1 in CM38 dataset

```{r identify_genes_MMP}

# #get the probesets from CM9 and CM38
# probeset_mmp <- MMP%>%
#   filter(abs(t_CM9) >= 1 &
#            abs(t_CM38) >= 1)%>%
#   pull(Probeset)

#get the probesets from CM38
probeset_mmp <- MMP%>%
  filter(abs(t_CM38) >= 1.5)%>%
  pull(Probeset)

```


```{r heatmap_MMP}


# Subset rmascreen data to the desired geneset 
rma_geneset <- rmascreen[rmascreen$Probeset %in% probeset_mmp,]

# Add the Gene Symbol to the geneset expression data table 
# so it can be used to label heatmap rows
rma_geneset <-  probeset %>%
	select(Probeset,Symbol) %>%
	inner_join(rma_geneset)

# scaled_mat will be the body of the heatmap of expression values
# Z-score the data
scaled_mat = t(scale(t(select(rma_geneset, -Probeset, -Symbol))))
rownames(scaled_mat) = rma_geneset$Symbol

# Provide column order as Ascending sPDL1
# This works because scaled_mat columns are in same order as SDRF rows
ordering = order(sdrfScreenSpdl1$PDL1_Av_Day1,
                 decreasing = TRUE)

# Top heatmap annotation is PDL1_Av_Day1 
ha_top = HeatmapAnnotation(barplot1 = anno_barplot(sdrfScreenSpdl1$PDL1_Av_Day1, 
												   axis = TRUE, 
												   baseline = 0,
						   						gp = gpar(fill = "black")),
						   annotation_height = unit(c(1), "cm"),
						   show_annotation_name = c(barplot1 = FALSE),
						 show_legend = FALSE)

#draw(ha_top,1:56)
# No Bottom heatmap annotation 

heatmap_object3 = Heatmap(scaled_mat,
						 col = colorRamp2(c(-1, 0, 1), c("dodgerblue", "white", "firebrick")), 
						 cluster_rows = FALSE,
						 km = 1,
						 row_names_gp = gpar(fontsize = 6),
						    				show_row_dend = FALSE,
						 show_row_names = TRUE,
						 row_names_side = "left",
						 cluster_columns = FALSE,
						 column_order = ordering,
						 column_title = "MMP genes that pass t>1.5",
						 column_title_gp = gpar(fontsize = 7),
						 show_column_names = FALSE,
						 width = unit(6, "cm"),
						 						show_heatmap_legend = FALSE,
						 height = unit(1, "cm"),
						 top_annotation = ha_top) 

# https://github.com/jokergoo/ComplexHeatmap/issues/110

gb_heatmap3 = grid.grabExpr(draw(heatmap_object3,
                                padding = unit(c(2, 10, 2, 2), "mm")))

#To use cowplot, convert the heatmap gtree returned from grid.grabExpr into a gtable:

heatmap_gtable_mmp <- gtable_matrix("heatmap_gtable", matrix(list(gb_heatmap3)), unit(1, "null"), unit(1, "null"))

```

\newpage
# Outputs

```{r cowplot_figure}

## Make the final pdf
fig_file <- paste(results_dir, "/Figure4_Affy_BL_CM38.pdf",
  sep="")

pdf(file=fig_file, width = 7.5, height = 9)

middle_right <-cowplot::plot_grid(heatmap_gtable2,
                   heatmap_gtable_mmp,
                   ncol=1,
                   nrow = 2,
                   scale = c(1,1),
                   rel_heights = c(1,0.8),
                   rel_widths = c(1),
                   labels = c("C","D"))

middle_row <-cowplot::plot_grid(barplot_nes,
                   middle_right,
                   ncol=2,
                   nrow = 1,
                   scale = c(1,1),
                   rel_heights = c(1),
                   rel_widths = c(1,1),
                   labels = c("B",""))

# bottom_row <-cowplot::plot_grid(heatmap_gtable_mmp,
#                    NULL,
#                    ncol=2,
#                    nrow = 1,
#                    scale = c(1,1),
#                    rel_heights = c(1),
#                    rel_widths = c(1,1),
#                    labels = c("D",""))


cowplot::plot_grid(heatmap_gtable_cs,
                   middle_row,
                   ncol=1,
                   nrow = 3,
                   scale = c(1,1),
                   rel_heights = c(1,0.9, 0.5),
                   rel_widths = c(1),
                   labels = c("A",""))

dev.off()

```

```{r cowplot_Supplementary}

## Make the final pdf
fig_file <- paste(results_dir, "/Figure4S_Affy_BL_CM38.pdf",
  sep="")

pdf(file=fig_file, width = 7.5, height = 9)

cowplot::plot_grid(barplot_baseline_myeloid_38,
                   barplot_baseline_neut_38,
                   barplot_nes_all, 
                   ncol=2,
                   nrow = 3,
                   scale = c(1,1,1),
                   rel_heights = c(1,1.5,0.8),
                   rel_widths = c(1),
                   labels = c("a","b","c",""))

dev.off()

```
