---
title: "Assemble the Supplementary tables"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clear the environment
rm(list = ls())
# Free up memory by forcing garbage collection
invisible(gc())  
# Manually set the seed to an arbitrary number for consistency in reports
myseed <- 9
##_ Set knitr root directory to correspond to project working directory 
##_  setting based on structure with code in <project dir>/code
knitr::opts_knit$set(root.dir = here::here())
```

# Paths, Packages, Functions, Palettes

```{r Paths}

# Provide paths
data_dir <- "./Data"
results_dir <- "./Outputs/Manuscript"

```


```{r LoadPackages}
library(knitr) # report generation


library(openxlsx)

##_ alternatively load tidyverse which bundles: ggplot2, dplyr, tidyr, readr, purrr and tibble
##_ use suppressPackageStartup() to hide package load messages
suppressPackageStartupMessages(library(tidyverse))  
```
\newpage
# Objective

Build Supplementary Tables


## Resources
IMM150 signatures:
+ Angio: VEGFA, KDR, ESM1, PECAM1, ANGPTL4, and CD34; 
+ Teff: CD8A, EOMES, PRF1, IFNG, and CD274; 
+ myeloid inflammation: IL-6, CXCL1, CXCL2, CXCL3, CXCL8, and PTGS2. 

## Data Sources

```{r loaddata}

#One row per patient sPDL1 data
spdl1_file <- paste(data_dir, "sPDL1_CM9_CM38_Annotated_v2.txt", sep = "/")
spdl1 <- read_tsv(spdl1_file)


# SDRF (Sample and Data Relationship Format) file from Array Express with fa01 update 
#CM009_annotation_of_Affymetrix_sPDL1.rmd
sdrf_file_9 <- paste(data_dir, "E-MTAB-3218 sdrf_response_sPDL1_Sigs_CS.Rds", sep = "/")
sdrf_9 <- readRDS(sdrf_file_9)

# SDRF (Sample and Data Relationship Format) file from Array Express with FA01 update
#FA01 = final analysis lock 01
#CM0038_annotation_of_Affymetrix_sPDL1.rmd
sdrf_file_38 <- paste(data_dir, "submission2439_annotare_2019sdrf_sPDL1_Sigs_CS.Rds", sep = "/")
sdrf_38 <- readRDS(sdrf_file_38)


# # Affymetrix probeset to Gene annotation
probeset_all_file <- paste(data_dir, "probeset_annotation.rds", sep = "/")
probeset_all <- readRDS(probeset_all_file)

# Affymetrix geneset probeset to Gene annotation
probeset_file <- paste(data_dir, "Probeset_RCC_Genesets.txt", sep = "/")
probeset <- read_tsv(probeset_file)

#GSEA_Hallmark results for both trials
GSEA_file <- paste(data_dir, "GSEA_Hallmark_CM9_CM38.txt", sep = "/")
GSEA <- read_tsv(GSEA_file)

#MMP results for both trials
MMP_file <- paste(data_dir, "MMPs_CM9_CM38_BLlimmaResults.txt", sep = "/")
MMP <- read_tsv(MMP_file)

```

CheckMate 009 Affymetrix sample annotation tied to individual (Sample and Data Relationship Format, SDRF file) created in CM009_annotation_of_Affymetrix_sPDL1.rmd was loaded from:

+ *`r sdrf_file_9`*

CheckMate 038 Affymetrix sample annotation tied to individual (Sample and Data Relationship Format, SDRF file) created in CM038_annotation_of_Affymetrix_sPDL1.rmd was loaded from:

+ *`r sdrf_file_38`*



```{r screen_sdrf_datasets}

# Baseline Subject sample 
sdrfScreenSpdl1_9 <- sdrf_9 %>%
	filter(biopsy.timepoint == "Screen" &
	         !is.na(PDL1_log2_Day1))

# Baseline Subject sample 
sdrfScreenSpdl1_38 <- sdrf_38 %>%
	filter(sampling_time_point == "screen" &
	         !is.na(PDL1_log2_Day1))


```

\newpage

# Table of Genesets used

Goal is to list the genes used for signature analysis.
In supplemental table I manually added MMPs to table S2

```{r genesets}

##### Add CytolyticScore

cyt_probesets <- c("5551_at",
         "3001_at")

probeset$Geneset_Cytolytic <- NA
probeset$Geneset_Cytolytic[probeset$Probeset %in% cyt_probesets] <- "Geneset_Cytolytic"

#### Add MMP genes
## This doesn't work because "Probeset_RCC_Genesets.txt" Doesn't have all the MMP genes
# mmp_probesets <- MMP%>%
#   filter(!is.na(Probeset))%>%
#   pull(Probeset)
# 
# probeset$Geneset_MMP60 <- NA
# probeset$Geneset_MMP60[probeset$Probeset %in% mmp_probesets] <- "Geneset_MMP60"


## keep columns used
columnsUsed <- c("Probeset",
                 "ENTREZ",
                 "LocusLink",
                 "Symbol",
                 "IRIS_Most_Specific",
                 "Geneset_ccRCC",
                 "Geneset_IMmotion150",
                 "Geneset_Merck18",
                 "Geneset_EMTstroma",
                 "Geneset_Javelin",
                 "Geneset_Cytolytic")

probeset <- probeset%>%
  select(one_of(columnsUsed))

# remove rows where all 6 geneset columns are NA
probesetsUsed <- probeset[rowSums(is.na(probeset[, c(6:11)])) != 6,]

rm(probeset)

```

#GSEA sets

Goal is to take the 6 shared+same direction HALLMARK pathways and identify the genes from the Core enrichment list for CM9 that are shared with CM38. 


```{r GSEA_sets}

shared_hallmark <- GSEA%>%
  filter(Compare_Direction == "Same" &
           Dataset == "CM38")%>%
  pull(Name)

probeset <- probeset_all%>%
  select("Probeset",
                 "ENTREZ",
                 "LocusLink",
                 "Symbol")

  for(i in shared_hallmark){

#https://stackoverflow.com/questions/42867124/comma-separated-string-split
#b <- unlist(strsplit(a, ","))

#get the leading edge genes from CM9
genes_9 <- GSEA$core_enrichment[which(GSEA$Name == i &
                                        GSEA$Dataset == "CM9")]
genes_9 <-  unlist(strsplit(genes_9, "/"))
  
#get the leading edge genes from CM38
genes_38 <- GSEA$core_enrichment[which(GSEA$Name == i &
                                        GSEA$Dataset == "CM38")]
genes_38 <-  unlist(strsplit(genes_38, "/"))

#find overlap
genes_shared <- genes_9[genes_9%in%genes_38]

#get probesets for shared leading edge "CoreEnrichment"
probeset_shared <- probeset_all$Probeset[probeset$Symbol%in%genes_shared]

probeset$foo <- NA

probeset$foo[probeset$Probeset %in% probeset_shared] <- "CoreEnrichment_CM9_CM38"

probeset <- rename(probeset,
                 !!i := foo)

}

# remove rows where all 6 HALLMARK columns are NA
probesetsShared <- probeset[rowSums(is.na(probeset[, c(5:10)])) != 6,]

rm(probeset, probeset_all)

```



# Correlation analyses

## CM9 Correlation: baseline sPDL1 to baseline numeric scores

From CIBERSORT, I am using  pre-treatment columns where more than 20% of samples have a non-zero value.

```{r CM9_correlation_baseline_baseline}

#here are all the pretreatment cibersort columns
#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1_9%>%
  select(starts_with("CS_pre"))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c("IM150_Angio.Score",
		   "IM150_MyeloidInfl.Score",
		   "IM150_Teff.Score",
		   "TIS",
    "Javelin.Score",
	"EMTstroma.Score",
	cibersort_columns)

cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#complete sPDL1 (same # in this case but not for CM38)
ScreenSpdl1 <- sdrfScreenSpdl1_9%>%
  filter(!is.na(PDL1_Av_Day1))

#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

correlation <- cor.test(ScreenSpdl1$PDL1_log2_Day1, y = ScreenSpdl1[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}

corr_9_bl_bl  <- df_data

kable(arrange(corr_9_bl_bl, p.value_pearson),
      title = "CM9: Correlation for Baseline sPDL1 versus Affymetrix scores")

```


## CM9 Correlation: change sPDL1 to baseline numeric scores

From CIBERSORT, I am using  pre-treatment columns where more than 20% of samples have a non-zero value.

There are 3 outliers for change in spdl1: 9159.603, 1887.669, -1943.826. I removed with a filter.

```{r CM9_correlation_baselineAffy_changeSPDL1}

#here are all the pretreatment cibersort columns
#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1_9%>%
  select(starts_with("CS_pre"))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c("IM150_Angio.Score",
		   "IM150_MyeloidInfl.Score",
		   "IM150_Teff.Score",
		   "TIS",
    "Javelin.Score",
	"EMTstroma.Score",
	cibersort_columns)

cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#patients with sPDL1 change
ScreenSpdl1 <- sdrfScreenSpdl1_9%>%
  filter(!is.na(PDL1_DIFF_Day29),
         PDL1_DIFF_Day29 <1800,
         PDL1_DIFF_Day29 >-1900)


#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

correlation <- cor.test(ScreenSpdl1$PDL1_DIFF_Day29, y = ScreenSpdl1[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}
 
corr_9_change_bl  <- df_data

kable(arrange(corr_9_change_bl, p.value_pearson),
      title = "CM9: Correlation for Change in sPDL1 versus baseline Affymetrix scores")

```



## CM9 Correlation: change sPDL1 to change CS

For CIBERSORT DIFF, I am using only columns where more than 20% of the Affy 42 sample pairs have a non-zero value.

I also explicitly remove the samples with DIFF = 0, which are samples with CS = zero before and after - even a tiny value produces a non-zero value for DIFF.

There are 3 outliers for D29 change in sPDL1: 9159.603, 1887.669, -1943.826 pg/ml. I removed with a filter.(only one is actually present in this subset so 38->37 samples)

```{r CM9_correlation_change_change}

#here are all the DIFF cibersort columns
#Just the 42 Affy-paired samples
mat_cs <- sdrfScreenSpdl1_9%>%
  filter(Subject_Affy_Status == "42Pairs")%>%
  select(starts_with("CS_DIFF"))
  
#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value
#with 37 samples this means >7 samples have to have a value!
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c(cibersort_columns)

cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#get paired samples with diff in sPDL1  = 38 pts
diffSpdl1 <- sdrfScreenSpdl1_9%>%
  filter(!is.na(PDL1_DIFF_Day29),
                Subject_Affy_Status == "42Pairs",
         PDL1_DIFF_Day29 <1800,
         PDL1_DIFF_Day29 >-1900)

#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

#explicitly remove the zero value diffs, which are samples with CS = zero before and after
#https://stackoverflow.com/questions/49786597/r-dplyr-filter-with-a-dynamic-variable-name
nonzero <- diffSpdl1%>%
        filter(!!rlang::sym(i) != 0)  
  
correlation <- cor.test(nonzero$PDL1_DIFF_Day29, y = nonzero[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}

corr_9_change_change  <- df_data 

kable(arrange(corr_9_change_change, p.value_pearson),
      title = "CM9: Correlation for change in sPDL1 versus CS_DIFF")

```



## CM38 Correlation: baseline sPDL1 to baseline numeric scores

From CIBERSORT, I am using  pre-treatment columns where more than 20% of samples have a non-zero value.

```{r CM38_correlation_baseline_baseline}

#here are all the pretreatment cibersort columns
#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1_38%>%
  select(starts_with("CS_pre"))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c("IM150_Angio.Score",
		   "IM150_MyeloidInfl.Score",
		   "TIS",
	"EMTstroma.Score",
	"Cytolytic.Score",
	cibersort_columns)


cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#complete sPDL1
ScreenSpdl1 <- sdrfScreenSpdl1_38%>%
  filter(!is.na(PDL1_Av_Day1))

#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

correlation <- cor.test(ScreenSpdl1$PDL1_log2_Day1, y = ScreenSpdl1[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}
 
corr_38_bl_bl  <- df_data 

kable(arrange(corr_38_bl_bl, p.value_pearson),
      title = "CM38: Correlation for Baseline sPDL1 versus Affymetrix scores")

```


## CM38 Correlation: change sPDL1 to baseline numeric scores

From CIBERSORT, I am using pre-treatment columns where **more than 10% of samples** have a non-zero value.

There are 3 outliers that affect correlation in sdrf_38$PDL1_DIFF_Day29:  5280.128,1769.973069, -3692.529

(only 5280,1769 have Affy)

So 35 samples go down to 33 samples without outliers.

```{r CM38_correlation_baselineAffy_changeSPDL1}

#here are all the pretreatment cibersort columns
#here are all the cibersort columns
mat_cs <- sdrfScreenSpdl1_38%>%
  select(starts_with("CS_pre"))

#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.90)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c("IM150_Angio.Score",
		   "IM150_MyeloidInfl.Score",
		   "TIS",
	"EMTstroma.Score",
	"Cytolytic.Score",
	cibersort_columns)

cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#patients with sPDL1 change
#remove 3 outliers
ScreenSpdl1 <- sdrfScreenSpdl1_38%>%
  filter(!is.na(PDL1_DIFF_Day29),
         PDL1_DIFF_Day29 <1700,
         PDL1_DIFF_Day29 >-1700)


#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

correlation <- cor.test(ScreenSpdl1$PDL1_DIFF_Day29, y = ScreenSpdl1[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}

corr_38_change_bl  <- df_data 

kable(arrange(corr_38_change_bl, p.value_pearson),
      title = "CM38: Correlation for Change in sPDL1 versus baseline Affymetrix scores")

```


## CM38 Correlation: change sPDL1 to change CS

For CIBERSORT, I am using only columns where more than 20% of the Affy 42 sample pairs have a non-zero value.

I also explicitly remove the samples with DIFF = 0, which are samples with CS = zero before and after - even a tiny value produces a non-zero value for DIFF.

There is are two outliers that affect correlation in sdrf_38$PDL1_DIFF_Day29:  5280.128,1769.973069, -3692.529

(only 5280,1769 have Affy)



```{r CM38 correlation_change_change}

#here are all the DIFF cibersort columns
#Just the 36 Affy-paired samples
mat_cs <- sdrfScreenSpdl1_38%>%
  filter(SubjectArrayCount == "2")%>%
  select(starts_with("CS_DIFF"))
  
#count the percentage of zero values for each cell type/column
percent_zero <-lapply(mat_cs, function(x){ length(which(x==0))/length(x)})

#find the columns where more than 20% of samples have a non-zero value
selectPct <- percent_zero[sapply(percent_zero, function(x) x < 0.80)]

## Get names of columns where more than 20% of samples have a value 
cibersort_columns <-names(selectPct)

#Define list to test
comp_scores <- c(cibersort_columns)

cor_methods <- c("pearson",
                 "spearman",
                 "kendall")

#get paired samples with diff in sPDL1  =  pts
diffSpdl1 <- sdrfScreenSpdl1_38%>%
  filter(!is.na(PDL1_DIFF_Day29),
                SubjectArrayCount == "2",
         PDL1_DIFF_Day29 <1700,
         PDL1_DIFF_Day29 >-1700)


#initialize results table
df_run <- NULL

#make a table to 'left_join' the results
df_data <- data.frame("Biomarker" = comp_scores)

#loop to run every score through every correlation
for(j in cor_methods){
  
  for(i in comp_scores){

#explicitly remove the zero value diffs, which are samples with CS = zero before and after
#https://stackoverflow.com/questions/49786597/r-dplyr-filter-with-a-dynamic-variable-name
nonzero <- diffSpdl1%>%
        filter(!!rlang::sym(i) != 0)  
  
correlation <- cor.test(nonzero$PDL1_DIFF_Day29, y = nonzero[[i]],
                    method = j)
  
 
p.value <- round(correlation$p.value[1], 3)  
estimate <- round(correlation$estimate[[1]], 2)

df_run<- rbind(df_run, data.frame("Biomarker"=i, "p.value" = p.value, "Cor" = estimate))


}

#append the test onto the results column name
#https://stackoverflow.com/questions/45472480/how-to-rename-a-column-to-a-variable-name-in-a-tidyverse-way
  
pval_test <- paste0("p.value_", j)
cor_test <-  paste0("Cor_", j)

df_run <- rename(df_run,
                 !!pval_test := p.value)

df_run <- rename(df_run,
                 !!cor_test := Cor)

#add  results for test j
df_data <- left_join(df_data, df_run,
                     by = "Biomarker")

df_run <- NULL
  
}
  
corr_38_change_change <- df_data

kable(arrange(df_data, p.value_pearson),
      title = "CM38: Correlation for change in sPDL1 versus CS_DIFF")

```


# Outputs

```{r spreadsheet}


require(openxlsx)

# list_of_datasets <- list("Name of DataSheet1" = dataframe1,
#                          "Name of Datasheet2" = dataframe2)
# 
# write.xlsx(list_of_datasets, file = "writeXLSX2.xlsx")

xlsx_file <- paste(results_dir,
                   "TABLE_Supplementary_Tables.xlsx",
                   sep="/")


list_of_datasets <- list("TS1_sPDL1_values" =spdl1,
                        "TS2_Genesets" = probesetsUsed,
                         "TS3_CM9_Corr_bl_bl" = corr_9_bl_bl,
                         "TS4_CM38_Corr_bl_bl" = corr_38_bl_bl,
                         "TS5_HALLMARK_GSEA_result" = GSEA,
                         "TS6_HALLMARK_GSEA_shared" = probesetsShared,
                         "TS7_MMP_CM9_CM38_bl" = MMP,
                         "TS8_CM9_Corr_chg_chg" = corr_9_change_change,
                         "TS9_CM38_Corr_chg_chg" = corr_38_change_change,
                         "TS10_CM9_Corr_chg_bl" = corr_9_change_bl,
                         "TS11_CM38_Corr_chg_bl" = corr_38_change_bl)

write.xlsx(list_of_datasets, file = xlsx_file, append=TRUE)


```



